On the presentation of orthogonal groups over polynomial rings

### 2019-05-20 dl-paper.tex
Minor

\DeclareMathOperator{\Ker}{Ker}
\DeclareMathOperator{\St}{St}
\DeclareMathOperator{\G}{G}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\HH}{H}
\DeclareMathOperator{\K}{K}
\DeclareMathOperator{\GW}{GW}
\DeclareMathOperator{\colim}{colim}
\newcommand{\inv}{^{-1}}

@@

\subsection{The subgroups $P_\alpha(0)$, $P_\alpha(*)$ and their properties.} Throughout this subsection $\Phi$ is a simply-laced root system of rank $\geq 3$ and $\alpha$ denotes some fixed root of $\Phi$.
\begin{df} \label{defP0}
Denote by $P_\alpha(0)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by the images of the following elements of $\St(\Phi, A[X])$ under the
 map induced by the embedding $A[X] \subseteq A[X, X^{-1}]$:
\begin{enumerate}
 \item $z_{\beta}(Xf, \xi), \beta \in \Phi \text{ such that }\alpha + \beta \in \Phi\text{ or } \alpha - \beta \in \Phi;$
 \item $z_{\beta}(f, X\xi),\ \beta \in \Phi \text{ such that }\alpha - \beta \in \Phi;$
 \item $z_{\beta}(f, \xi),\ \beta \in Z_0(\alpha);$
 \item $x_{-\alpha}(X^2f);$
 \item $x_{\alpha}(f)$.
\end{enumerate}
In the above formulae and for the rest of this subsection $f \in M[X]$ and $\xi \in A[X]$. \end{df}

@@

Denote by $P_\alpha(*)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by $P_\alpha(0)$ and the elements $x_{-\alpha}(Xm)$, $m \in M$. The following lemma shows that the resulting subgroup is sufficiently large.
\begin{lemma} \label{Pstar-large} The subgroup $P_\alpha(*)$ contains the image of $K(A[X], M[X])$ in $\St(\Phi, A[X, X^{-1}])$. \end{lemma}
\begin{proof} Clearly, $P_\alpha(*)$ contains the elements $x_\beta(Xf)$ for all $\beta \in \Phi$ and $z_\beta(Xf, \xi)$ for $\beta \in \Phi \setminus \{\pm \alpha\}$, $f \in M[X]$, $\xi\in A[X]$ hence by~\cref{thm:Stepanov} $P_\alpha(*)$ contains whole $\St(\Phi, A[X], XM[X])$. The required assertion now follows from~\cref{Kgen-strong} and the preceding remark. \end{proof}

The above lemma allows us to characterize $P_\alpha(*)$ as follows: it consists precisely of those elements $g \in \St(\Phi, A[X], M[X])$ for which $ev_{X=0}^*(g)$ lies in the subgroup generated by
 elements $x_{\beta}(m),\ \beta \in Z_+(\alpha)$ and $z_{\gamma}(m, \zeta),\ \gamma \in Z_0(\alpha),$ where $m \in M$, $\zeta \in A$ 
 and $Z_+(\alpha)$, $Z_0(\alpha)$ are as in~\cite[Proposition~5.7]{St71}.

@@

\[ \xymatrix{ \overline{\St}(\Phi, A, M) \ar[r]^{j_+i_+} \ar[d]^{i^-} & \overline{G}_M^{\geq 0} \ar@{^{(}->}[d] \\
              G^{\leq 0} \ar[r]^{j_-} & \St(\Phi, A[X, X^{-1}]), } \] 
Denote by $\overline{G}_M^0$ the image of the left arrow $i_-$ in $G^{\leq 0}$ and by $C$ the set of left cosets of $\overline{G}_M^0$ in $G^{\leq 0}$, i.\,e.
$C = \{ g\overline{G}^0_M \mid g \in G^{\leq 0} \}.$

Denote by $V$ the subset of $\St(\Phi, A[X, X^{-1}]) \times G^{\leq 0} \times (1 + M)^*$ consisting of those triples $(g, h, u)$ for which 
$\pi(g, h, u) = g \cdot j_-(h) \cdot \{ X, u \} \in \overline{G}_M^{\geq 0}$.
Denote by $\overline{V}$ the image of $V$ in $\St(\Phi, A[X, X^{-1}]) \times C \times (1 + M)^*$
under the obvious map induced by the projection $G^{\leq 0} \to C$
 (of course, these triples still can be characterized by the property that $(g, h \overline{G}^0_M, u) \in \overline{V}$ iff $\pi(g, h, u) \in \overline{G}_M^{\geq 0}$).

The next step of the proof is to define for $\alpha \in \Phi$, $a \in A$ a partial function $T_\alpha(aX^{-1}) \colon V \not\to V$.
This function is defined for those triples $(g, h, u)$ for which $\pi(g, h, u) \in P_\alpha(*) \subseteq \overline{G}_M^{\geq 0}$.
If $\pi(g, h, u) \in P_\alpha(m)$ for some $m \in M$, then $T_\alpha(aX^{-1})$ is given by the following identity:
\[ T_\alpha(aX^{-1}) (g, h, u) = \left( x_\alpha(aX^{-1})\cdot g ,\
                                         h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right),\ 
                                         u \cdot (1 + am)\right).\]
It is easy to verify that $\pi(T_\alpha(aX^{-1}) (g, h, u)) = S_\alpha(a, \pi(g, h, u)).$

\begin{lemma} $T_\alpha(aX^{-1})$ descends to a well-defined total function $T_\alpha(aX^{-1}) \colon \overline{V} \to \overline{V}$. \end{lemma}
\begin{proof}
 First of all, let us show that the resulting function is total. 
 We need to show that for every $(g, h\overline{G}^0_M, u) \in \overline{V}$ we can find a representative $h' \in h\overline{G}^0_M$
 such that $\pi(g, h', u) \in P_\alpha(*)$. Set $h' = h \cdot ev_{X=0}(\pi(g, h, u))^{-1}$, then
  \[\pi(g, h', u) = g \cdot j_-(h) \cdot \{ X, u \} \cdot \left(ev_{X=0}(g \cdot j_-(h))\right)^{-1} \in Im(K(A[X], M[X]) \to \St(\Phi, A[X, X^{-1}])) \subseteq P_\alpha(*).\]

 Next, we have to show that the value of $T_\alpha(aX^{-1})$ is well-defined. Suppose $(g, h_1, u), (g, h_2, u) \in V$ are such that 
  both $\pi(g, h_1, u)$ and $\pi(g, h_2, u)$ lie in $P_\alpha(*)$. Consequently, $j_-(h_1 h_2^{-1}) \in P_\alpha(*)$.

\end{proof}

>>>>

\DeclareMathOperator{\Ker}{Ker}
\DeclareMathOperator{\Img}{Im}
\DeclareMathOperator{\St}{St}
\DeclareMathOperator{\G}{G}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\HH}{H}
\DeclareMathOperator{\K}{K}
\DeclareMathOperator{\GW}{GW}
\DeclareMathOperator{\colim}{colim}
\newcommand{\inv}{^{-1}}

@@

\subsection{The subgroups $P_\alpha(0)$, $P_\alpha(*)$ and their properties.} Throughout this subsection $\Phi$ is a simply-laced root system of rank $\geq 3$ and $\alpha$ denotes some fixed root of $\Phi$.
\begin{df} \label{defP0}
Denote by $P_\alpha(0)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by the images of the following elements of $\St(\Phi, A[X])$ under the
 map $j_+ \colon \overline{\St}(\Phi, A[X], M[X]) \to \St(\Phi, A[X, X^{-1}])$:
\begin{enumerate}
 \item $z_{\beta}(Xf, \xi), \beta \in \Phi \text{ such that }\alpha + \beta \in \Phi\text{ or } \alpha - \beta \in \Phi;$
 \item $z_{\beta}(f, X\xi),\ \beta \in \Phi \text{ such that }\alpha - \beta \in \Phi;$
 \item $z_{\beta}(f, \xi),\ \beta \in Z_0(\alpha);$
 \item $x_{-\alpha}(X^2f);$
 \item $x_{\alpha}(f)$.
\end{enumerate}
In the above formulae and for the rest of this subsection $f \in M[X]$ and $\xi \in A[X]$. \end{df}

@@

Denote by $P_\alpha(*)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by $P_\alpha(0)$ and the elements $x_{-\alpha}(Xm)$, $m \in M$. The following lemma shows that the resulting subgroup is sufficiently large.
\begin{lemma} \label{Pstar-large} The subgroup $P_\alpha(*)$ contains the image of $K(A[X], M[X])$ in $\St(\Phi, A[X, X^{-1}])$. \end{lemma}
\begin{proof} Clearly, $P_\alpha(*)$ contains the elements $x_\beta(Xf)$ for all $\beta \in \Phi$ and $z_\beta(Xf, \xi)$ for $\beta \in \Phi \setminus \{\pm \alpha\}$, $f \in M[X]$, $\xi\in A[X]$ hence by~\cref{thm:Stepanov} $P_\alpha(*)$ contains whole $j_+(\overline{\St}(\Phi, A[X], XM[X]))$. The required assertion now follows from~\cref{Kgen-strong} and the preceding remark. \end{proof}

\begin{rem} \label{Pstar-char} The above lemma allows us to characterize $P_\alpha(*)$ as follows: it consists of images under $j_+$ of those elements 
 $g \in \overline{\St}(\Phi, A[X], M[X])$ for which $ev_{X=0}^*(g)$ lies in the subgroup $C_\alpha(A, M) := \Img(\overline{\St}(Z_0(\alpha), A, M) \to \overline{\St}(\Phi, A, M)) \ltimes \hat{U}(Z_+(\alpha), M)$ 
 (i.\,e. the subgroup generated by elements $x_{\beta}(m),\ \beta \in Z_+(\alpha)$ and $z_{\gamma}(m, \zeta),\ \gamma \in Z_0(\alpha),$ where $m \in M$, $\zeta \in A$ 
 and $Z_+(\alpha)$, $Z_0(\alpha)$ are as in~\cite[Proposition~5.7]{St71}). \end{rem}

@@

\[ \xymatrix{ \overline{\St}(\Phi, A, M) \ar[r]^{j_+i_+} \ar[d]^{i^-} & \overline{G}_M^{\geq 0} \ar@{^{(}->}[d] \\
              G^{\leq 0} \ar[r]^{j_-} & \St(\Phi, A[X, X^{-1}]), } \] 
Denote by $\overline{G}_M^0$ the image of the left arrow $i_-$ in $G^{\leq 0}$.

Denote by $V$ the subset of $\St(\Phi, A[X, X^{-1}]) \times G^{\leq 0} \times (1 + M)^*$ consisting of those triples $(g, h, u)$ for which 
$p(g, h, u) := g \cdot j_-(h) \cdot \{ X, u \}$ belongs to $\overline{G}_M^{\geq 0}$.
We let $\overline{G}_M^0$ act on $V$ on the right by $(g, h, u) \cdot h_0 = (g, h \cdot h_0, u)$, $h_0 \in \overline{G}_M^0$.
It is clear that this action is well defined. Denote by $\overline{V}$ the quotient set $V / \overline{G}_M^0$.

The next step of the proof is to define for $\alpha \in \Phi$, $a \in A$ a partial function $T_\alpha(aX^{-1}) \colon V \not\to V$.
This function is defined for those triples $(g, h, u)$ for which $p(g, h, u) \in P_\alpha(*) \subseteq \overline{G}_M^{\geq 0}$.
If $p(g, h, u) \in P_\alpha(m)$ for some $m \in M$, then $T_\alpha(aX^{-1})$ is given by the following identity:
\[ T_\alpha(aX^{-1}) (g, h, u) = \left( x_\alpha(aX^{-1})\cdot g ,\ h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right),\ u \cdot (1 + am)\right).\]
It is easy to verify that $p(T_\alpha(aX^{-1}) (g, h, u)) = S_\alpha(a, p(g, h, u)).$

\begin{lemma} $T_\alpha(aX^{-1})$ descends to a well-defined total function $T_\alpha(aX^{-1}) \colon \overline{V} \to \overline{V}$. \end{lemma}
\begin{proof}
 First of all, let us show that the resulting function is total. 
 Fix $c = [(g, h, u)] \in \overline{V}$. Since $p(g, h, u) \in \overline{G}^{\geq 0}_M$ there exists $g_1 \in \overline{\St}(\Phi, A[X], M[X])$ such that $j_+(g_1) = p(g, h, u).$
 Set $h_0 = ev_{X=0}(g_1))^{-1}$, then, clearly, $g_1 \cdot i_+(h_0) \in K(A[X], M[X])$ and 
 $p(g, h \cdot i_-(h_0), u) = j_+(g_1) \cdot j_-(i_-(h_0)) = j_+(g_1 \cdot i_+(h_0)) \in j_+(K(A[X], M[X])) \subseteq P_\alpha(*)$,
  which shows that $(g, h \cdot i_+(h_0), u)$ is the representative of $c$ on which $T_\alpha(aX^{-1})$ is defined.

 Next, let us show that the value of $T_\alpha(aX^{-1})$ does not depend on the choice of a representative.
 Let $(g, h_1, u), (g, h_2, u)$ be a pair of representatives of the same $c \in \overline{V}$ for which both $p(g, h_1, u)$ and $p(g, h_2, u)$ lie in $P_\alpha(*)$.
 By definition, $h_1 h_2^{-1} = i_-(h_0)$, for $h_0 \in \overline{\St}(\Phi, A, M)$, moreover, $p(g, h_1, u) \cdot p(g, h_2, u)^{-1} = j_- i_-(h_0) = j_+ i_+(h_0) \in P_\alpha(*)$,
 Clearly, $ev_{X=1}(j_+ i_+ (h_0)) = ev_{X=1}(i_+(h_0)) = ev_{X=0}(i_+(h_0)) = h_0$.
 On the other hand, by~\cref{Pstar-char} there exists $g_1 \in \overline{\St}(\Phi, A[X], M[X])$ such that $ev_{X=0}(g_1) \in C_\alpha(A, M)$
  and $j_+(g_1) = j_+ i_+ (h_0)$.
 Now we have: $h_0 = ev_{X=1}(j_+(g_1)) = ev_{X=1}(g_1)$

 %Invoking injective stability for $\K_1(\Phi, -)$ and surjective stability $\K_2(\Phi, -)$ we can find $h_0 \in \overline{\St}(Z_0(\alpha), A, M) \ltimes U(Z_+(\alpha), M)$ such that $\pi(h_0) = \pi(j_-(h_1 h_2^{-1}))$.
\end{proof}
